---
title: "Lab 12 - Smoking during pregnancy"
author: "Thomas Huang"
date: "2026-04-13"
output: github_document
---

# This is my 11th lab. But I choose Lab 12 as my 11th lab.

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse)
library(infer)
library(openintro)
```

### Exercise 1

Numeric variables are fage, mage, weeks, visits, gained, and weight.
```{r}
data(ncbirths)

glimpse(ncbirths)

hist(ncbirths$fage)
hist(ncbirths$mage)
hist(ncbirths$weeks)
hist(ncbirths$visits) #Seems to be an outlier / outliers.
hist(ncbirths$gained)
hist(ncbirths$weight)
```

### Exercise 2

```{r}
ncbirths_white <- ncbirths %>% 
  filter(whitemom == "white")
```

### Exercise 3

The observations should be mostly independent of each other because a mother is unlikely to give birth to multiple babies in 2004.
The sample size is 714. It's reasonably large.
The shape of the distribution looks normal.
```{r}
glimpse(ncbirths_white)
hist(ncbirths_white$weight)
```

### Exercise 4

```{r}
set.seed(123)

# Generate bootstrapped mean
ncbirths_white_boot <- ncbirths_white %>%
  specify(response = weight) %>% 
  generate(reps = 15000, type = "bootstrap") %>% 
  calculate(stat = "mean")

# Recenter the mean
ncbirths_white_boot$stat_centered <- ncbirths_white_boot$stat - mean(ncbirths_white_boot$stat, na.rm = TRUE) + 7.43

# Figure
ggplot(ncbirths_white_boot, aes(x = stat_centered)) +
  geom_histogram(fill = "skyblue", color = "black") +
  geom_vline(xintercept = 7.43, linetype = "dashed", color = "red")

# Two-tailed p
F <- ecdf(ncbirths_white_boot$stat_centered)
p_val <- 2*min(F(7.43), 1-F(7.43))
p_val

# Interpretation
# Birth weight has not changed since 1995, p = .992.
```

### Exercise 5

Weight distribution is more centered and compact among nonsmokers. Both distributions are left skewed.
```{r}
ncbirths_r <- ncbirths %>% 
   filter(!is.na(habit))

ggplot(ncbirths_r, aes(x = weight)) +
  geom_histogram() +
  facet_wrap(~habit, ncol = 2) 
```

### Exercise 6

First, group summaries for an NA group don't make sense. Second, NAs in weight may cause errors when calculating group summaries (which needs something like na.omit() to fix).
```{r}
ncbirths_clean <- ncbirths %>% 
   filter(!is.na(habit), !is.na(weight))
```

### Exercise 7

This difference itself is not meaningful. It is not enough for inference.
```{r}
diff <- ncbirths_clean %>%
  group_by(habit) %>%
  summarize(mean_weight = mean(weight)) %>%
  summarize(diff = mean_weight[habit == "nonsmoker"] -
                   mean_weight[habit == "smoker"]) %>% 
  pull(diff)
```

### Exercise 8
$H_0$: The average birth weights of the two groups are the same. Alternatively, $\mu_{Smoker}=\mu_{NonSmoker}$.
$H_1$: The average birth weights of the two groups are different. Alternatively, $\mu_{Smoker} \neq \mu_{NonSmoker}$.

### Exercise 9

Permutation is my choice. This time I'm comparing the esimates of two groups by first assuming the effect is 0 (null). Permutation better fits this idea.
```{r}
set.seed(123)

diffs <- replicate(5000, {
  shuffled <- ncbirths_clean %>%
    mutate(habit = sample(habit))  
  
  shuffled %>%
    group_by(habit) %>%
    summarize(m = mean(weight), .groups = "drop") %>%
    summarize(perm.diff = m[habit == "nonsmoker"] -
                     m[habit == "smoker"]) %>%
    pull(perm.diff)
})

F <- ecdf(diffs)
p <- 2 * min(F(diff), 1 - F(diff))
p
# p = .0336. The smnoker group is associated with lower average birth weight.
```

### Exercise 10

# The 95% CI does not include 0. This indicates the difference between the two groups is statistically significant.
```{r}
diff_boot <- ncbirths_clean %>%
  specify(response = weight, explanatory = habit) %>% 
  generate(reps = 15000, type = "bootstrap") %>% 
  calculate(stat = "diff in means", order = c("nonsmoker", "smoker"))

get_confidence_interval(diff_boot, level = 0.95, type = "percentile")
```





